---
title: "Multiple Regression 3"
subtitle: Sam Larralde and Ambrosio Rivera
format: pdf
editor: visual
---

# **Activity 11: Statistical reasoning 3: multiple regression DAG's**

Install packages

Load packages

```{r}
library(brms) # for statistics
library(tidyverse) # for data wrangling

# a function to scale and center. from rethinking package
standardize <- function(x) {
    x <- scale(x)
    z <- as.numeric(x)
    attr(z,"scaled:center") <- attr(x,"scaled:center")
    attr(z,"scaled:scale") <- attr(x,"scaled:scale")
    return(z)
}
```

# **1. DAG practice**

# Q 1.1 Make DAG

dag { bb="0,0,1,1" A \[pos="0.436,0.218"\] B \[pos="0.434,0.488"\] C \[pos="0.788,0.372"\] U \[pos="0.142,0.360"\] X \[pos="0.160,0.701"\] Y \[pos="0.794,0.691"\] A -\> C A -\> U C -\> B C -\> Y U -\> B U -\> X X -\> Y }

# Q 1.2 Identifying forks

X \<- U -\> B

U \<-A -\> C

B \<- C -\> Y

# Q 1.3 Identify colliders

U → B ← C

C → Y ← X

# Q 1.4 Modify the DAG

dag { bb="0,0,1,1" A \[pos="0.436,0.218"\] B \[pos="0.434,0.488"\] C \[pos="0.788,0.372"\] U \[pos="0.142,0.360"\] V \[pos="0.656,0.526"\] X \[pos="0.160,0.701"\] Y \[pos="0.794,0.691"\] A -\> C A -\> U C -\> B C -\> Y U -\> B U -\> X V -\> C V -\> Y X -\> Y }

# Q 1.5 Identify paths

x-\> y

U -\> X -\> Y

A -\> U -\> X -\> Y

# Q 1.6 Identify open backdoor paths

U -\> X

C -\> Y

V -\> Y

# Q 1.7 Identify variables to close the back doors

U, C, Y

# **Foxes: Regression Practice informed by DAG's**

Load library

```{r}
# Load in the fox data
foxes <- read.csv('https://raw.githubusercontent.com/rmcelreath/rethinking/refs/heads/master/data/foxes.csv', sep = ';')
```

check out fox data

```{r}
# Check out the fox data
?foxes
head(foxes)
```

# Q 2.1 Identify the fundamental relations in the fox DAG

fork: groupsize \<-avgfood -\> weight

colider: avgfood -\> weight \<- groupsize

pipe: area -\> avgfood -\> weight

pipe: avgfood-\>groupsize -\> weight

Total causal influence of area on weight

```{r}
fox_dat <- foxes %>%
  as_tibble() %>%
  select(area, avgfood, weight, groupsize) %>%
  mutate(across(everything(), standardize))
```

simulate from some priors

```{r}
n <- 1000
priorsims <- tibble(group = seq_len(n),
       alpha = rnorm(n, 0, 0.2), # prior for alpha
       beta = rnorm(n, 0, 2)) %>% # prior for beta
  expand(nesting(group, alpha, beta), # the expand function gives us all possible combinations of the arguments
         area = seq(from = -2, to = 2, length.out = 100)) %>% # set up a range of areas
  mutate(weight = alpha + beta * area) # calculate weight from the parameters and area
```

make a plot

```{r}
ggplot(priorsims, aes(x = area, y = weight, group = group)) +
  geom_line(alpha = 1 / 10) +
  labs(x = "Standardized Area", y = "Standardized Weight")
```

# Q 2.2 Minimum fox weight

We think at minimum the fox might weigh 10 kg

# Q 2.3 Maximum fox weight

We think a reasonable maximum weight for a fox is 14g

# Q 2.4 Modify simulation plot

min= 10kg, max = 14kg

```{r}
ggplot(priorsims, aes(x = area, y = weight, group = group)) +
  geom_line(alpha = 1 / 10) +
  labs(x = "Standardized Area", y = "Standardized Weight")
  geom_hline(y= 10)
  geom_hline(y=40)
```

```{r}
mean_w <- mean(foxes$weight)
sd_w   <- sd(foxes$weight)
min_std <- (10 - mean_w) / sd_w
max_std <- (14 - mean_w) / sd_w
ggplot(priorsims, aes(x = area, y = weight, group = group)) +
  geom_line(alpha = 1/10) +
  geom_hline(yintercept = min_std, color = "red", linetype = "dashed") +
  geom_hline(yintercept = max_std, color = "red", linetype = "dashed") +
  labs(x = "Standardized Area",
       y = "Standardized Weight")
```

# Q 2.5 Evaluate prior predictive simulation

The priors do not seem reasonable, as weight will not cross into negative values.

# Q 2.6 Refine priors

```{r}
n <- 1000
priorsims <- tibble(group = seq_len(n),
       alpha = rnorm(n, 6, 0.5),# prior for alpha
       beta = rnorm(n, 0,.2)) %>% # prior for beta
  expand(nesting(group, alpha, beta), # the expand function gives us all possible combinations of the arguments
         area = seq(from = -2, to = 2, length.out = 100)) %>% # set up a range of areas
  mutate(weight = alpha + beta * area) # calculate weight from the parameters and area
```

```{r}
mean_w <- mean(foxes$weight)
sd_w   <- sd(foxes$weight)
min_std <- (10 - mean_w) / sd_w
max_std <- (14 - mean_w) / sd_w
ggplot(priorsims, aes(x = area, y = weight, group = group)) +
  geom_line(alpha = 1/10) +
  geom_hline(yintercept = min_std, color = "red", linetype = "dashed") +
  geom_hline(yintercept = max_std, color = "red", linetype = "dashed") +
  labs(x = "Standardized Area",
       y = "Standardized Weight")
```

Run Models

```{r}
food_on_area <- brm(avgfood ~ 1 + area, 
                    data = fox_dat, 
                    family = gaussian,
                    # Here we set the priors that we investigated earlier
                    prior = c(prior(normal(6, 0.5), class = Intercept),
                              prior(normal(0,.2), class = b,),
                              prior(exponential(1), class = sigma)),
                    iter = 4000, warmup = 2000, chains = 4, cores = 4, seed = 1234,
                    file = "output/food_on_area")
```

```{r}
summary(food_on_area)
```

# Q 2.7 Run a model for the impact of food on fox weight

```{r}
# fox weight model
m.fox.weights <- 
  brm(data = fox_dat, # Give it fox data
      # Choose a gaussian (normal) distribution
      family = gaussian,
      # Specify the model here. 
      weight ~ 1 + avgfood,
      # Here's where you specify parameters for executing the Markov chains
      # We're using similar to the defaults, except we set cores to 4 so the analysis runs faster than the default of 1
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      # Setting the "seed" determines which random numbers will get sampled.
      # In this case, it makes the randomness of the Markov chain runs reproducible 
      # (so that both of us get the exact same results when running the model)
      seed = 4,
      # Save the fitted model object as output - helpful for reloading in the output later
      file = "output/m.fox.weight")
```

```{r}
summary(m.fox.weights)
```

The slope shows a negative correlation between food and fox weight. This unexpected so something should be missing or mistaken.

# Q 2.8 Is there a variable we should condition upon?

We should condition group size and area.

# Q 2.9 What is your hypothesis about how group size affects weight?

We hypothesize that group size is positively correlated with average weight. We think that larger groups will have a larger territory and more food available to them and therefore foxes from larger groups could weight more. Less individual resources more competition when they are in a bigger group. But its a trade off that they can have larger territory and more protection.

Now lets run the model
```{r}
group_on_weight <- brm(weight ~ 1 + groupsize, 
                       data = fox_dat, 
                       family = gaussian,
                       prior = c(prior(normal(0, 0.2), class = Intercept),
                                 prior(normal(0, 0.5), class = b,),
                                 prior(exponential(1), class = sigma)),
                       iter = 4000, warmup = 2000, chains = 4, cores = 4, seed = 1234,
                       file = "output/group_on_weight")
```


```{r}
summary(group_on_weight)
```
Let's add in groupsize to block the pipe weight->groupsize->avgfood:
```{r}
food_direct <- brm(weight ~ 1 + avgfood + groupsize, 
                   data = fox_dat,
                   family = gaussian,
                   prior = c(prior(normal(0, 0.2), class = Intercept),
                             prior(normal(0, 0.5), class = b,),
                             prior(exponential(1), class = sigma)),
                   iter = 4000, warmup = 2000, chains = 4, cores = 4, seed = 1234,
                   file = "output/food_direct")

summary(food_direct)
```


# Q 2.10a What are the effects of avgfood and groupsize now that you have accounted for both variables?

We can conclude our hypothesis, that average food is positively correlated with weight, and large groups make smaller foxes. The slope for average food was positive, meaning when food increased so did weight. While the slope for group size was negative so larger groups means less weight.

# Q 2.10b How does this interpretation change your interpretation from the univariate regressions of each variable separately with weight?

Initially, the older model suggested that more food meant lighter foxes but the newer model shows that to be in reverse.

# Q 2.10c Provide a small discussion (2-4 sentences) explaining in your own words why these results turned out the way they did, in the context of the ecological system of fox territories. Include why you think that the univariate regressions may have suggested no relationship while the multiple regression suggests a different answer

Larger groups of foxes have more competition between each other for resources, which will lower their food availabity and weight. While more food on its own of course leads to individually heavier foxes. The model may have been misinterpreting food availability before, as fox group size has an impact on the other variables.


